---
title: "ggplot 2"
output: github_document
---

```{r, setup}
library(tidyverse)
library(patchwork)
library(viridis)
```


## Load the weather data 

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),#pick out 2 datasets
    var = c("PRCP", "TMIN", "TMAX"), #download from the internet
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

## Start with a familiar one

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)
```
sometime people might not know what's the meaning of the x and y

## Labels(best to do when show sb else)

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Mininum daily temperature(c)",
    y = "Maximum daily temperature(c)",##add label to the whole plot and xy axis
    caption = "Data from rnoaa package; temperatures in 2017"#add caption(zhushi)
  )# make it easier for other people to use
```

## Scales (eg. change x = tmax to x = log(tmax)...)

start with the same plot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Mininum daily temperature(c)",
    y = "Maximum daily temperature(c)",##add label to the whole plot and xy axis
    caption = "Data from rnoaa package; temperatures in 2017"#add caption(zhushi)
  ) +
  scale_x_continuous( #set the plot first, then begin to scale
    breaks = c(-15,0,15),#change numbers showed on x axis
    labels = c("-15 c","0","15")#change the name of the 3 points on x axis we set before
  ) +
  scale_y_continuous(
    trans = "sqrt",
    position = "right"#change the position of the y axis
  )
```

Look at color scales

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Mininum daily temperature(c)",
    y = "Maximum daily temperature(c)",##add label to the whole plot and xy axis
    caption = "Data from rnoaa package; temperatures in 2017"#add caption(zhushi)
  ) +
  scale_color_hue(#many other kinds of functions other than "hue"
    name = "location", 
    h = c(100,300))#set the color of the points
```


```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Mininum daily temperature(c)",
    y = "Maximum daily temperature(c)",##add label to the whole plot and xy axis
    caption = "Data from rnoaa package; temperatures in 2017"#add caption(zhushi)
  ) +
  viridis::scale_color_viridis(
    name = "location",
    discrete = TRUE)#tell it that is is a discrete varieble, or use "scale_color_viridis_d() insteading of "scale_color_viridis()"
```


## Themes 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Mininum daily temperature(c)",
    y = "Maximum daily temperature(c)",##add label to the whole plot and xy axis
    caption = "Data from rnoaa package; temperatures in 2017"#add caption(zhushi)
  ) +
  viridis::scale_color_viridis(
    name = "location",
    discrete = TRUE) +#tell it that is is a discrete varieble
  theme(legend.position = "bottom")#put legend in the bottom
theme_bw()
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Mininum daily temperature(c)",
    y = "Maximum daily temperature(c)",##add label to the whole plot and xy axis
    caption = "Data from rnoaa package; temperatures in 2017"#add caption(zhushi)
  ) +
  scale_color_viridis_d() +
theme_bw()+
  theme(legend.position = "bottom")#put legend in the bottom
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Mininum daily temperature(c)",
    y = "Maximum daily temperature(c)",##add label to the whole plot and xy axis
    caption = "Data from rnoaa package; temperatures in 2017"#add caption(zhushi)
  ) +
  scale_color_viridis_d() +
theme_minimal()+
  theme(legend.position = "bottom")#put legend in the bottom
```


```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Mininum daily temperature(c)",
    y = "Maximum daily temperature(c)",##add label to the whole plot and xy axis
    caption = "Data from rnoaa package; temperatures in 2017"#add caption(zhushi)
  ) +
  viridis::scale_color_viridis(
    name = "location",
    discrete = TRUE) +#tell it that is is a discrete varieble
 #***** ggthemes::thems_excel()
```

##setting options(set every time the same)

```{r, eval = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.comtinuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale
```


##data args in `geom`

```{r}
#create 2 sub dataframes
central_park_df = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

wakiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = wakiki, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park_df)
```

##`patchwork package`

remember faceting?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .5) +
  facet_grid(. ~ name)
```

what happens when you want multipanel plots but can't facet ...?

```{r}
tmax_tmin_p = 
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  theme(legend.position = "name")

prcp_dens_plot = 
  weather_df %>% 
  filter(prcp > 0) %>% #only keep the >0 part
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = .5)

tmax_data_plot = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, coor = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none")

(prcp_dens_plot + tmax_data_plot) / tmax_tmin_p
```

##data manipulation
control your factors.

```{r}
weather_df %>% 
  mutate(
    name = fct_reorder(name, tmax)
  ) %>% #tidy your data before making the plots
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_boxplot()
```

```{r}
weather_df %>% 
  mutate(
    name = factor(name),#change name from charecter to factor
    name = forcats::fct_relevel(name,c("Waikiki_HA"))
  ) %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = .5)##color is not needed here
```

What if I wanted densities for tmin and tmax simultaneously?

```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,#get 2 curves at the same time 
    names_to = "observation",
    values_to = "temperatures"
  ) %>% 
  ggplot(aes(x = temperatures, fill = observation)) +
  geom_density(alpha = .5) +
  facet_grid(. ~ name)
```

```{r}
pulse_df = 
  haven::read_sas("data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% ##many na
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi",
    names_prefix = "bdi_score_"
  ) %>% 
  mutate(visit = recode(visit, "bl" = "00m")) 

pulse_df %>% 
  ggplot(aes(x = visit, y = bdi)) +
  geom_boxplot()

pulse_df %>% #keep track each person's change
  ggplot(aes(x = visit, y = bdi)) +
  geom_point(size = .2)+
  geom_line(aes(group = id))
```


## revisit the pups

Data from the FAS study.

```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv") %>% 
  janitor::clean_names() %>% 
  mutate(sex = recode(sex, `1` = "male", `2` = "female"))

litters_data = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  seperate(group, into = c("dose", "day_of_tx"),sep = 3)

fas_data = left_join(pup_data, litters_data, by = "litter_number")

fas_data %>% 
  ggplot(aes(x = dose, y = pd_ears)) +
  geom_violin() +
  facet_grid(. ~ day_of_tx ~ outcome)
```

