---
title: "ggplot 1"
output: github_document
---

```{r, setup}
library(tidyverse)
library(ggridges)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6
,
out.width = "90%"
)
```


## Load the weather data 

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),#pick out 2 datasets
    var = c("PRCP", "TMIN", "TMAX"), #download from the internet
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

## Scatterplots

tmax vs tmin

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()#relationship between tmin and tmax
```

you can save ggplots

```{r}
ggp_tmax_tmin = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_tmax_tmin
#ggsave()
```

create my first scatterplot ever

```{r}
ggplot(weather_df,aes(x = tmin, y = tmax)) +
  geom_point()
```

New approach, same plot

```{r}
tail(weather_df)
```

```{r}
weather_df %>% #add sth like filter or mutate is ok
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

##Opyions to read_csv

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
```

Save and edit a plot object

```{r}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))#save the plot in the df

weather_plot = geom_point()
```


## Advanced scatterplot

Start with the same one and make it fancy

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE)#se = limit
```

What about the `aes` placement..?


```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .3) +#alpha = Í¸Ã÷¶È
  geom_smooth(se = FALSE) + #color only apply to point, and smooth will not get the color from point
  facet_grid(. ~ name)#devide 3 name dataframe into 3 separate form
```

Let's facet somenthings!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .2,size = .3) +#size is the size of the points
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)#dot means don't create rows
```

let's make one more scatterplot.

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, size = prcp)) +
  geom_point(alpha = .3) +
  facet_grid(. ~ name) +
  geom_smooth(se = FALSE)
```

##use data manipulation as part of this

```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(
    tmax = tmax * (9 / 5) + 32,
    tmin = tmin * (9 / 5) + 32
  ) %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex() 
```

Let's  combine some elements and try a new plot

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) + 
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```


## somr small notes

How many geoms have to exist?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```

You can use a neat geom!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_density2d() +
  geom_point(alpha = .3)
```

## Univariate plots(only have 1 varie)

Histograms are really great

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

can we add color ..

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +#color = color is only on the outline, fill = color also filled in
  geom_histogram(position = "dodge")
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +#color = color is only on the outline, fill = color also filled in
  geom_histogram()+
  facet_grid(. ~name)
```

Let's try a new geometry!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .3, adjust = .5)
```

What about box plots??

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot()
```

Trendy plots :-)

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin() + 
  stat_summary()# also show mean
```

Ridge plots -- the most popular plot of 2017

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = name)) +
geom_density_ridges()
```

## Save and Embed

Let's save a scatterplot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax,color = name)) +
  geom_point(alpha = .5)

```

What about embedding...

```{r}
weather_plot
```

embed at different size
```{r, fig.width=12, fig.height=6}
weather_plot
```